\documentclass[12pt, a4paper]{article}

\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{parskip}
\usepackage{times}
\usepackage{ragged2e}

\pagestyle{empty}

\begin{document}

% ── Sender block ──────────────────────────────────────────────────────────────
\noindent
\textbf{Moise Iradukunda Ingabire}\\
MSIT (Applied AI \& Cybersecurity)\\
Carnegie Mellon University Africa\\
KG 590 St, Kigali, Rwanda\\
\href{mailto:mii@andrew.cmu.edu}{mii@andrew.cmu.edu}\\
May 21, 2026

\vspace{0.4cm}

% ── Recipient block ───────────────────────────────────────────────────────────
\noindent
The Editorial Board, \textit{Future Internet}\\
MDPI AG\\
St. Alban-Anlage 66, 4052 Basel, Switzerland

\vspace{0.5cm}

\noindent Dear Editor,

\vspace{0.3cm}

% ── Body ──────────────────────────────────────────────────────────────────────
We are pleased to submit the \textbf{second revised manuscript} of
\textbf{``AI-Augmented Compliance Auditing for Cloud Systems: A Hybrid ML--LLM Approach''}
(Manuscript ID: futureinternet-4270373), in response to the Round~2 reviewer comments.

We are grateful to all three reviewers for their continued engagement. Reviewer~3 has
confirmed that the manuscript is now \textit{``suitable for publication''} and requested
no further changes. Reviewer~1 raised two minor issues, both of which have been fully
addressed as described below.

\vspace{0.4cm}
\noindent\textbf{\large Round 2 Changes (this submission)}

\vspace{0.2cm}
\noindent\textbf{Issue 1: Residual old numbers (99.99\%, 99.88\%, 100\%):}
Reviewer~1 noted that several legacy figures appeared to remain in the revised manuscript.
We have reviewed every occurrence and taken targeted action:

\begin{itemize}
  \item \textbf{99.99\% (7 occurrences):} All appear in explicit retraction context
        (e.g., ``the previously reported 99.99\% F1, attributable to an 86.3\% data-leakage
        rate, is retracted'') and are intentionally retained for transparency.

  \item \textbf{99.88\% (2 occurrences, fixed):} This fine-tuning result (SSH-domain
        only) now carries an explicit scope qualifier in both the Phase~3 narrative
        (Section~3.5) and the RQ summary table (Table~9):
        \textit{``(on the same SSH domain; this result reflects SSH-specific vocabulary
        adaptation and should not be interpreted as cross-domain or general-purpose
        performance)''}.

  \item \textbf{100\% occurrences:} All remaining 100\% figures are either (a) the 100\%
        total in the dataset composition table; (b) per-family F1 values for IA and MP
        families, each accompanied by Wilson 95\% CIs; (c) adversarial detection rates for
        T1190, T1110, and T1078, each accompanied by Wilson 95\% CIs; or (d) the 100\% FPR
        for OOV Windows/HTTP log types, explained as expected architectural behaviour.
        None represent inflated performance claims.
\end{itemize}

\noindent\textbf{Issue 2: LLM sample size and generalization claims:}
Reviewer~1 confirmed the evaluation is acceptable for exploratory/applied evaluation but
noted it does not support broad claims of LLM superiority. Language has been updated in
three targeted locations:

\begin{itemize}
  \item \textbf{Phase~3 description (Section~3.5):} ``a direct solution to the
        generalization gap'' replaced with ``preliminary evidence that semantic LLM
        reasoning can address the vocabulary-based generalization gap at acceptable
        deployment cost, while remaining insufficient for broad claims of LLM superiority
        in cybersecurity auditing without larger-scale validation.''

  \item \textbf{Finding F4 (Section~4.3):} ``consistently superior to XGBoost zero-shot
        across all categories'' replaced with ``consistently higher than XGBoost zero-shot
        across all evaluated categories; this constitutes preliminary applied validation
        and does not support broad claims of LLM superiority in general cybersecurity
        auditing contexts.''

  \item \textbf{Statistical Scope paragraph (Section~4.3):} New sentence added:
        ``These sample sizes support exploratory and applied evaluation conclusions; they
        are not sufficient to support broad claims of LLM superiority in general
        cybersecurity auditing, and results should be interpreted as preliminary evidence
        for the specific log types and compliance context evaluated.''
\end{itemize}

\vspace{0.4cm}
\noindent\rule{\textwidth}{0.4pt}
\vspace{0.2cm}

\noindent\textbf{\large Round 1 Changes (first major revision, May 9, 2026, for reference)}

\vspace{0.2cm}
\noindent\textbf{The most consequential change in Round~1} was the full identification and
correction of a critical data leakage issue. A systematic audit of the original training
corpus found that \textbf{86.3\%} of synthetic training rows contained embedded compliance
labels in the \texttt{log\_message} field, causing the classifier to memorize label strings
rather than learn security semantics. This produced the previously reported 99.99\% F1
score, which was explicitly retracted. The synthetic data generator was rewritten from
scratch, and the corrected \textbf{XGBoost v2 achieves 85.45\% macro-F1}
(Wilson 95\% CI: [84.9\%, 86.0\%]; 5-fold CV F1: 85.20\% $\pm$ 0.25\%) on a
verified leakage-free corpus of 150{,}000 training rows.

Additional Round~1 changes:

\begin{itemize}
  \item \textbf{Statistical rigor}: Wilson 95\% confidence intervals reported for every
        result in every table. All 100\% detection rates accompanied by Wilson CIs.

  \item \textbf{LLM evaluation expanded}: $n$=200 per log type ($n$=628 total; HTTP
        limited to $n$=28 real SecRepo Squid proxy logs). GPT-4o-mini macro accuracy:
        \textbf{92.3\%} across four log types (37.5\% real enterprise data). Llama-3.2-3B
        removed as underpowered.

  \item \textbf{Adversarial validation expanded}: 2 to \textbf{5 MITRE ATT\&CK v14
        scenarios} (T1059.001, T1190, T1110, T1078, T1562.001), macro detection rate
        \textbf{92.8\%}, 0.0\% FPR on real SSH/PAM logs ($n$=75).

  \item \textbf{Standardized joint evaluation}: New Hybrid System Combined Evaluation
        (Section~4.4) on identical 180 real-world logs. XGBoost alone: 62.4\% F1 (54.5\%
        FPR); GPT-4o-mini alone: 90.8\% F1 (6.4\% FPR); Hybrid: \textbf{85.1\% F1
        (6.4\% FPR)}.

  \item \textbf{Theoretical formulation clarified}: All variables in
        $C(k) = \alpha D_k + \beta V_k + \gamma R_k$ explicitly defined with default
        weights and experimental grounding. Output dimensionality corrected to 169 controls.

  \item \textbf{Overstated claims moderated}: ``First'', ``proves'', ``strictly more
        expressive'', and cost-efficiency comparisons softened or removed.

  \item \textbf{Reproducibility}: Full supplementary appendix documenting dataset
        provenance, hyperparameter tables, and phase-by-phase experimental logs.
\end{itemize}

\vspace{0.4cm}
\noindent\rule{\textwidth}{0.4pt}
\vspace{0.2cm}

No new experiments, tables, or structural sections were added in Round~2. All Round~2
changes are targeted text edits at the three locations noted above, consistent with the
minimal intervention requested by Reviewer~1.

The revised manuscript remains original, is not under review elsewhere, uses only
publicly available datasets, and complies fully with MDPI's ethical guidelines. The
Article Processing Charge is understood and accepted.

We are grateful for the thorough engagement from all three reviewers across both rounds
of review, and we hope the manuscript now meets the publication standards of
\textit{Future Internet}.

\vspace{0.5cm}

\noindent Sincerely,

\vspace{0.8cm}

\noindent\textbf{Corresponding Author:} Prof. Jema David Ndibwile\\
\noindent\href{mailto:jndibwil@andrew.cmu.edu}{jndibwil@andrew.cmu.edu}\\[0.3cm]
\noindent\textbf{First Author:} Moise Iradukunda Ingabire\\
\noindent\href{mailto:mii@andrew.cmu.edu}{mii@andrew.cmu.edu}

\end{document}
