\documentclass[11pt,conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{float}
\usepackage{balance}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usetikzlibrary{shapes,arrows,positioning,calc,fit,backgrounds}
\usepackage{needspace}

\lstset{
    basicstyle=\scriptsize\ttfamily,
    breaklines=true,
    frame=single,
    numbers=none,
    language=Python,
    showstringspaces=false,
    float=!htbp
}

\begin{document}

\title{An AI-Augmented Unified Compliance Auditor for Hybrid Cloud Environments: Integrating Adversarial Security Techniques for Rwanda's Regulatory Landscape}

\author{\IEEEauthorblockN{Moise IRADUKUNDA INGABIRE}
\IEEEauthorblockA{\textit{College of Engineering} \\
\textit{Carnegie Mellon University Africa}\\
Kigali, Rwanda \\
mii@andrew.cmu.edu}}

\maketitle

\begin{abstract}
This research presents the design, implementation, and validation of an AI-augmented unified compliance auditor specifically tailored for Rwanda's National Cyber Security Authority (NCSA) Minimum Cybersecurity Standards. The system employs an evolving hybrid architecture: initially XGBoost multi-label classification for evidence-to-control mapping, subsequently enhanced with Model Context Protocol (MCP) integration enabling Large Language Model (LLM) semantic analysis achieving 100\% accuracy on event-based compliance classification. Through systematic analysis of adversarial techniques including genetic algorithms, dynamic programming for payload generation, AV/EDR evasion methods, and reverse shell exploits, we demonstrate how understanding offensive security informs defensive compliance validation. Our seven-engine microservices architecture, deployed on Kubernetes, implements \textbf{143 controls} (expanded from initial 47) with 96 system-auditable controls supported by 60 specialized evidence parsers. The system achieves 99.99\% F1-score ($\pm$0.01\%, 5-fold CV) on synthetic data, 99.88\% F1-score on real SecRepo authentication logs after minimal fine-tuning, and 100\% accuracy with semantic LLM analysis. The hybrid dataset strategy (70\% real-world public logs, 30\% Rwanda-specific synthetic data) addresses overfitting challenges inherent in purely synthetic training. The evolution from keyword-based (XGBoost) to semantic (LLM) classification eliminates the 7.98\% zero-shot limitation, enabling true generalization to novel log formats. This work represents the first ML-driven compliance system tailored to African regulatory frameworks, with resource-efficient design suitable for resource-constrained environments. Validation testing on Kubernetes demonstrates production readiness, providing a replicable blueprint for automated compliance auditing across African nations.
\end{abstract}

\begin{IEEEkeywords}
Compliance Auditing, Rwanda NCSA, Machine Learning, XGBoost, Large Language Models, Model Context Protocol, Multi-label Classification, Hybrid Cloud Security, NIST SP 800-53, Adversarial Security, Kubernetes
\end{IEEEkeywords}

\section{Introduction}

\subsection{Context and Motivation}

Enterprises and government agencies across Africa are undergoing rapid digital transformation, increasingly adopting hybrid cloud infrastructures. However, compliance and cybersecurity assurance remain fragmented and heavily manual, consuming up to 40\% of IT security budgets \cite{ponemon_cost}, often requiring more than 1,000 hours annually \cite{cogr_report}, and introducing error-prone, siloed approaches.

At a continental level, the African Union's Malabo Convention aims to create a unified legal framework for cybersecurity \cite{malabo2014}. However, its implementation has been slow, with only 15 of 54 African nations having ratified the convention as of 2024 \cite{au_cybersecurity_status}. This regulatory gap has prompted individual nations to take the lead. Rwanda stands out as a proactive leader, having established comprehensive NCSA Minimum Cybersecurity Standards in 2023 \cite{ncsa2023} and a National Cybersecurity Strategy for 2024-2029 \cite{rwanda_strategy_2024}, derived from the Malabo Convention and aligned with NIST SP 800-53 \cite{nist2020}.

Simultaneously, the cybersecurity threat landscape has evolved dramatically. Modern attacks employ sophisticated evasion techniques leveraging genetic algorithms \cite{ga_xss_paper}, dynamic programming for payload optimization \cite{dp_payload_paper}, and advanced AV/EDR bypass methods \cite{reverse_shell_paper, mitre_attack_sok}. Compliance frameworks must validate not merely the \textit{presence} of controls but their \textit{effectiveness} against such adversarial techniques.

\subsection{Research Problem}

Current compliance auditing approaches face three critical challenges:

\textbf{1. Manual and Resource-Intensive}: Traditional audits require extensive human effort, making continuous compliance monitoring infeasible for resource-constrained Rwandan institutions. Organizations conducting 3-5 internal audits annually achieve per-capita compliance costs averaging \$154, while those without internal audits face costs of \$341 \cite{ponemon_cost}.

\textbf{2. Checkbox Compliance}: Existing frameworks verify control presence (e.g., ``Is antivirus installed?'') without assessing effectiveness. Research demonstrates signature-based detection suffers 90-97\% evasion rates against modern techniques \cite{dp_payload_paper, reverse_shell_paper}.

\textbf{3. Binary Classification Inadequacy}: Compliance is inherently multi-dimensional---a single log event may provide evidence for multiple controls simultaneously. Binary classification (compliant/non-compliant) oversimplifies this reality.

\subsection{Research Contributions}

This work makes four novel contributions:

\begin{enumerate}
    \item \textbf{Adversarial-Aware Compliance Framework}: First integration of offensive security research (GA-based payload generation, AV/EDR evasion, reverse shell techniques) into defensive compliance validation methodology.

    \item \textbf{Multi-Label Evidence-to-Control Mapping}: Novel formulation treating compliance auditing as multi-label classification, where evidence maps to multiple controls (e.g., failed SSH login $\rightarrow$ AC-7, IA-2, AU-2, SI-4).

    \item \textbf{Rwanda-Specific ML Architecture}: First AI-driven compliance system tailored to African regulatory framework (NCSA 2023), optimized for resource constraints (137x smaller than BERT alternatives, 50x faster inference).

    \item \textbf{Production-Ready Microservices Deployment}: Seven-engine Kubernetes-deployed system with demonstrated scalability, validated through end-to-end testing and multi-pod deployment.
\end{enumerate}

\subsection{Research Questions}

\textbf{Primary Research Question}: Given the limitations of manual and tool-siloed auditing in Rwandan hybrid cloud environments, can an AI-augmented compliance auditor employing hybrid architecture (ML-based evidence routing + rule-based evaluation) for NIST SP 800-53 and Rwanda NCSA standards achieve 65-80\% macro F1-score for control classification and $\geq$50\% audit cycle reduction while maintaining explainability?

\textbf{Secondary Research Questions}:
\begin{enumerate}
    \item How do adversarial techniques (GA, DP, LLM-based payload generation, AV/EDR evasion) inform compliance control effectiveness validation?
    \item What are the comparative strengths of BERT, LSTM, and XGBoost for evidence-to-control mapping in resource-constrained environments?
    \item Can hybrid datasets (real-world public logs + Rwanda-specific synthetic data) overcome overfitting limitations of purely synthetic training?
    \item How does multi-label classification improve upon binary compliance classification for African regulatory contexts?
\end{enumerate}

\subsection{System Overview and Paper Organization}

We propose a seven-engine microservices architecture where logs flow through collection, analysis, decision-making, and reporting stages. Engine 1 (Log Collector) normalizes heterogeneous log sources; Engine 3 (MCP+LLM Analyzer \cite{mcp_anthropic}) performs hybrid rule-based and semantic evidence-to-control mapping; Engine 4 (Decision Engine) applies rule-based compliance evaluation per control; and Engine 6 (Report Generator) produces actionable compliance reports. Figure~\ref{fig:pipeline} illustrates the complete data flow.

The remainder of this paper is organized as follows: Section II reviews related work on adversarial security and AI-assisted compliance. Section III presents our methodology including system architecture and dataset strategy. Section IV details implementation. Section V presents evaluation results. Section VI discusses findings, theoretical contributions, ethical considerations, and limitations. Section VII outlines future work, and Section VIII concludes.

%==============================================================================
\section{Literature Review}

\subsection{Adversarial Security and Compliance}

Recent research demonstrates sophisticated attack automation through optimization techniques. Understanding these adversarial methods is critical for validating compliance control effectiveness.

\subsubsection{Genetic Algorithms for Security Testing}

Liu et al. \cite{ga_xss_paper} developed GAXSS, a genetic algorithm for XSS vulnerability detection achieving 100\% precision and 97.5\% accuracy. The DNA-encoded payload structure with multi-objective fitness function (execution success, tag closure, diversity, punctuation balance) demonstrates GA effectiveness for security testing. However, 81.5\% recall indicates 18.5\% of vulnerabilities remain undetected---a gap relevant for SI-10 (Input Validation) control validation.

Fine-tuning large language models for payload generation \cite{llm_payload_paper} achieved 94.73\% evasion rate against ClamAV through transfer learning and mutation strategies. This offensive capability informs defensive applications: compliance auditors can generate synthetic attack scenarios to validate SI-3 (Malicious Code Protection) control effectiveness.

Dynamic programming approaches \cite{dp_payload_paper} formulate payload obfuscation as Markov Decision Process optimization, achieving 97\% detectability reduction through systematic code transformations. The deterministic optimization enables reproducible adversarial testing for compliance validation.

\textbf{Research Gap}: None of these papers connect optimization techniques to regulatory compliance frameworks or demonstrate defensive applications for control effectiveness testing.

\subsubsection{AV/EDR Evasion Techniques}

Modern evasion techniques systematically bypass detection mechanisms. Custom reverse shell implementations \cite{reverse_shell_paper} demonstrate fileless execution reducing detection from 62.9\% (44/70 AV engines) to 1.4\% (1/70)---a 97\% evasion improvement through shellcode separation and memory-only execution.

The MITRE ATT\&CK framework documents 40+ defense evasion techniques under tactic TA0005 \cite{mitre_attack_sok}. A systematic review of 417 publications found that defense evasion tactics (T1027 Obfuscation, T1055 Process Injection, T1562 Impair Defenses) are the most commonly observed in real-world attacks \cite{mitre_attack_applications}.

\textbf{Compliance Implications}: These evasion techniques reveal that signature-based detection---still prevalent in Rwandan SMEs---provides false compliance assurance. Our compliance auditor addresses this by validating detection \textit{capability} through adversarial testing.

\subsubsection{African Cybersecurity Context}

Cybercrimes cost African economies an estimated \$3.5 billion annually, with Nigeria (\$649M), Kenya (\$210M), and South Africa (\$157M) bearing the heaviest losses \cite{serianu_africa}. Despite this, only 29 of 54 African countries have promulgated cybersecurity legislation \cite{itu_africa_cyber}.

Kenya's Computer Misuse \& Cybercrimes Act (2024) and Critical Information Infrastructure Regulations require continuous monitoring \cite{kenya_cyber_law}. Nigeria's Cybercrimes Act 2015 created a National Cybersecurity Fund \cite{nigeria_cyber}. Rwanda's approach is notable for establishing sector-specific minimum standards enforced through mandatory audits \cite{ncsa2023, rwanda_strategy_2024}.

\textbf{Research Gap}: No prior work maps evasion techniques to African compliance controls or provides frameworks for effectiveness-based validation in resource-constrained contexts.

\subsection{AI-Assisted Compliance and Risk Scoring}

Green \& Brown (2023) \cite{bert_audit} fine-tuned BERT for security audit log classification, achieving $>$93\% accuracy on benchmarks. Smith et al. (2022) \cite{lstm_audit} used LSTM autoencoders for anomaly detection, reporting 50\% audit cycle reduction. However, both studies assume well-resourced environments with GPU infrastructure---conditions not met in Rwanda's SME landscape.

Recent work on the SIEVE dataset \cite{sieve_dataset} demonstrates that SVM achieves macro-F1 of 0.93-0.97 on synthetic logs but degrades on real logs, while BERT shows similar behavior (0.95 on synthetic, 0.89 on real). This motivates our hybrid dataset approach.

\textbf{Our Contribution}: We demonstrate XGBoost achieves comparable performance (96\% on test logs) while being 137x smaller (3.2MB vs. 2.7GB BERT) and 50x faster ($<$1ms vs. 45ms), making it deployable on CPU-only infrastructure typical in African institutions.

\subsection{Multi-Label Classification}

Tsoumakas and Katakis \cite{tsoumakas2007} established foundational multi-label techniques. Our application to compliance auditing is novel: evidence inherently relates to multiple controls (e.g., ``Failed SSH login'' $\rightarrow$ AC-7, IA-2, AU-2, SI-4). Binary classification discards this multi-dimensional reality.

\textbf{Research Gap}: No prior compliance auditing system employs multi-label classification for evidence-to-control mapping, particularly in African regulatory contexts.

%==============================================================================
\section{Methodology}

\subsection{System Architecture}

\subsubsection{Seven-Engine Microservices Design}

The system employs a microservices architecture with seven independent engines, as shown in Figure~\ref{fig:architecture}.

\begin{figure}[!t]
\centering
\begin{tikzpicture}[
    node distance=0.8cm,
    engine/.style={rectangle, draw, fill=blue!20, text width=1.6cm, align=center, minimum height=0.7cm, font=\scriptsize, rounded corners=2pt},
    db/.style={cylinder, draw, fill=green!20, text width=1.4cm, align=center, minimum height=0.5cm, shape border rotate=90, font=\scriptsize},
    arrow/.style={->, >=stealth, thick}
]
% Engines
\node[engine] (e1) at (0,0) {E1: Log\\Collector};
\node[engine, right=0.6cm of e1] (e2) {E2: Doc\\Processor};
\node[engine, right=0.6cm of e2] (e3) {E3: MCP+LLM\\Analyzer};
\node[engine, below=0.8cm of e2] (e4) {E4: Decision\\Engine};
\node[engine, left=0.6cm of e4] (e6) {E6: Report\\Generator};
\node[engine, right=0.6cm of e4] (e5) {E5: Web UI};
\node[engine, below=0.8cm of e4] (e7) {E7: Auth};

% Databases
\node[db, below left=0.6cm and 0.3cm of e7] (pg) {PostgreSQL};
\node[db, below right=0.6cm and 0.3cm of e7] (redis) {Redis};

% Arrows
\draw[arrow] (e1) -- (e2);
\draw[arrow] (e2) -- (e3);
\draw[arrow] (e3) -- (e4);
\draw[arrow] (e4) -- (e6);
\draw[arrow] (e4) -- (e5);
\draw[arrow, dashed] (e7) -- (e4);
\draw[arrow, dashed] (e7) -- (e5);
\draw[arrow] (e4) -- (pg);
\draw[arrow] (e5) -- (redis);
\end{tikzpicture}
\caption{Seven-engine microservices architecture}
\label{fig:architecture}
\end{figure}

\begin{table}[!t]
\centering
\caption{Engine Functional Breakdown}
\label{tab:engines}
\begin{tabular}{@{}lp{5.5cm}@{}}
\toprule
\textbf{Engine} & \textbf{Function} \\
\midrule
Engine 1 & Log collection (syslog, Windows Events, files, APIs) \\
Engine 2 & Document processing (PDF evidence extraction) \\
Engine 3 & MCP+LLM hybrid analyzer with 60 evidence parsers and audit executor \\
Engine 4 & Decision engine (143 control-specific evaluators) \\
Engine 5 & Report generation (PDF/CSV compliance reports) \\
Engine 6 & Web UI (React dashboard, audit wizard, real-time progress, WebSocket) \\
Engine 7 & Authentication (JWT, RBAC) \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Architectural Rationale}:
\begin{itemize}
    \item \textbf{Separation of Concerns}: Engine 3's hybrid analyzer (rule-based fast path + LLM semantic analysis) routes evidence; Decision Engine evaluates compliance. This leverages both pattern matching for speed and LLM reasoning for semantic understanding. MCP's tool-use abstraction enables clean separation between compliance logic and AI provider APIs \cite{mcp_enterprise}.
    \item \textbf{Independent Scaling}: High-volume log collection can scale independently from computationally intensive LLM analysis.
    \item \textbf{Technology Heterogeneity}: Engines use optimal technologies (FastAPI, React, Claude/GPT APIs) without monolithic constraints. MCP standardizes the interface, enabling provider-agnostic LLM integration \cite{mcp_spec}.
\end{itemize}

\subsubsection{Data Flow}

The complete pipeline is illustrated in Figure~\ref{fig:pipeline}.

\begin{figure*}[!t]
\centering
\begin{tikzpicture}[
    node distance=0.9cm,
    source/.style={rectangle, draw, fill=blue!15, text width=1.6cm, align=center, minimum height=0.6cm, font=\scriptsize, rounded corners=2pt},
    process/.style={rectangle, draw, fill=orange!15, text width=1.8cm, align=center, minimum height=0.6cm, font=\scriptsize, rounded corners=2pt},
    output/.style={rectangle, draw, fill=green!15, text width=1.6cm, align=center, minimum height=0.6cm, font=\scriptsize, rounded corners=2pt},
    arrow/.style={->, >=stealth, thick}
]
% Sources
\node[source] (syslog) {Syslog};
\node[source, below=0.3cm of syslog] (windows) {Windows\\Events};
\node[source, below=0.3cm of windows] (docs) {Documents};

% Processing stages
\node[process, right=1.2cm of windows] (e1) {E1: Normalize\\+ Enrich};
\node[process, right=0.8cm of e1] (e3) {E3: MCP+LLM\\Analyze};
\node[process, right=0.8cm of e3] (e4) {E4: Decision\\Evaluate};
\node[process, right=0.8cm of e4] (e6) {E6: Generate\\Reports};

% Outputs
\node[output, right=0.8cm of e6] (pdf) {PDF Report};
\node[output, below=0.3cm of pdf] (dash) {Dashboard};

% Arrows from sources
\draw[arrow] (syslog.east) -- ++(0.3,0) |- (e1.west);
\draw[arrow] (windows) -- (e1);
\draw[arrow] (docs.east) -- ++(0.3,0) |- (e1.west);

% Processing flow
\draw[arrow] (e1) -- node[above, font=\tiny] {Evidence} (e3);
\draw[arrow] (e3) -- node[above, font=\tiny] {Control IDs} (e4);
\draw[arrow] (e4) -- node[above, font=\tiny] {Decisions} (e6);

% Output arrows
\draw[arrow] (e6) -- (pdf);
\draw[arrow] (e6.east) -- ++(0.2,0) |- (dash.west);

% Labels
\node[below=0.1cm of e3, font=\tiny, text width=2.2cm, align=center] {Hybrid: Rules\\+ LLM Semantic};
\end{tikzpicture}
\caption{End-to-end compliance auditing pipeline showing data flow from heterogeneous sources through hybrid MCP+LLM analysis, decision, and reporting stages}
\label{fig:pipeline}
\end{figure*}

For multi-label classification, each evidence item $x_i$ maps to a binary label vector:
\begin{equation}
\mathbf{y}_i = [y_{i1}, y_{i2}, \ldots, y_{i143}] \in \{0,1\}^{143}
\end{equation}
where $y_{ij} = 1$ if evidence $x_i$ relates to control $c_j$.

\subsection{Dataset Strategy}

\subsubsection{Evolution from Synthetic to Hybrid}

Initial purely synthetic data (100,000 events) revealed critical overfitting. All three baseline models achieved 100\% accuracy---a clear red flag. Investigation identified:

\textbf{Data Leakage}: The \texttt{status\_code} feature exhibited -0.97 correlation with the compliance label. Models memorized: \texttt{status\_code == 200} $\rightarrow$ compliant.

\textbf{Template Simplicity}: Synthetic generation used fixed templates lacking real-world complexity.

\subsubsection{Hybrid Dataset Composition}

To address overfitting, we integrated real-world public datasets as shown in Table~\ref{tab:dataset}.

\begin{table}[!t]
\centering
\caption{Hybrid Dataset Composition (200,000 logs)}
\label{tab:dataset}
\begin{tabular}{@{}lrrp{3cm}@{}}
\toprule
\textbf{Source} & \textbf{Count} & \textbf{\%} & \textbf{Type} \\
\midrule
NSL-KDD \cite{nslkdd} & 104,000 & 52\% & Network intrusions \\
LogHub \cite{loghub} & 36,000 & 18\% & System logs \\
Rwanda Synthetic & 60,000 & 30\% & NCSA-specific \\
\midrule
\textbf{Total} & \textbf{200,000} & \textbf{100\%} & Hybrid \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Rationale}: 70\% real-world data prevents overfitting to synthetic patterns while 30\% synthetic ensures coverage of Rwanda-specific NCSA requirements not present in public datasets.

\subsubsection{Additional Evaluation Datasets}

For cross-dataset validation, we evaluated on:
\begin{itemize}
    \item \textbf{SIEVE} \cite{sieve_dataset}: 30-class SIEM event classification dataset with expert-labeled categories
    \item \textbf{SecRepo Auth Logs} \cite{secrepo}: 86K real authentication logs with failed SSH attempts
    \item \textbf{LANL Authentication} \cite{lanl_auth}: Sampled 50K from 708M enterprise authentication events
\end{itemize}

\subsection{Model Development and Selection}

\subsubsection{Baseline Models}

Three models were trained for comparative analysis:

\textbf{BERT (110M parameters)}: Fine-tuned bert-base-uncased with frozen embeddings and first 10 layers. Training: 2 epochs, batch 16, learning rate 2e-5. Result: 12 hours training, 2.7GB model, 45ms latency.

\textbf{LSTM (758K parameters)}: 2-layer bidirectional LSTM, 128 hidden units, 0.3 dropout. Training: 5 epochs, vocabulary 946. Result: 2 hours training, 800MB model, 12ms latency.

\textbf{XGBoost}: Gradient boosting trees, max depth 6, learning rate 0.1. Training: 500 estimators (early stopping at 93). Result: 8 minutes training, 350MB model, $<$1ms latency.

\subsubsection{Production Selection: XGBoost}

XGBoost was selected based on resource efficiency critical for Rwanda (Table~\ref{tab:model_comparison}).

\begin{table}[!t]
\centering
\caption{Model Comparison for Production Deployment}
\label{tab:model_comparison}
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Metric} & \textbf{BERT} & \textbf{LSTM} & \textbf{XGBoost} \\
\midrule
Training Time & 12 hrs & 2 hrs & \textbf{8 min} \\
Model Size & 2.7 GB & 800 MB & \textbf{350 MB} \\
Latency & 45 ms & 12 ms & \textbf{$<$1 ms} \\
Memory & 2.1 GB & 580 MB & \textbf{380 MB} \\
Explainability & Low & Low & \textbf{High (SHAP)} \\
GPU Required & Yes & No & \textbf{No} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Feature Engineering}

\textbf{Numeric Features (5)}: Hour of day (0-23), day of week (0-6), business hours flag, network port, event severity.

\textbf{Text Features (25 TF-IDF)}: Vectorized log message field with unigrams, min document frequency 5, capturing terms like \textit{failed}, \textit{unauthorized}, \textit{blocked}, \textit{encrypted}.

Note: \texttt{status\_code} was removed from final model due to identified leakage.

\subsection{Multi-Label Formulation}

\textbf{Problem}: Map evidence $x_i$ to subset of controls $\mathcal{C}_i \subseteq \{c_1, \ldots, c_{143}\}$.

\textbf{Approach}: Multi-output XGBoost using scikit-learn's \texttt{MultiOutputClassifier} wrapper with base XGBoost estimators (100 trees, depth 6, learning rate 0.1), producing 143-dimensional output vectors corresponding to all implemented controls.

\textbf{Evaluation Metrics}:
\begin{align}
F1_{\text{macro}} &= \frac{1}{143}\sum_{k=1}^{143} F1_k \\
F1_{\text{micro}} &= \frac{2 \cdot P_{\text{all}} \cdot R_{\text{all}}}{P_{\text{all}} + R_{\text{all}}} \\
\text{Hamming Loss} &= \frac{1}{N \cdot 143}\sum_{i,j} \mathbb{1}(y_{ij} \neq \hat{y}_{ij})
\end{align}

\subsection{Evolution to Semantic LLM Analysis}

While XGBoost achieves excellent performance on in-distribution data, the 7.98\% zero-shot transfer limitation reveals a fundamental constraint: TF-IDF features capture vocabulary patterns rather than semantic meaning. A log containing ``authentication failure'' is correctly classified only if similar vocabulary appeared in training data.

\subsubsection{Model Context Protocol Integration}

To overcome vocabulary dependency, we developed a hybrid architecture integrating Large Language Models (LLMs) via Model Context Protocol (MCP) \cite{mcp_anthropic, mcp_spec}---an open standard introduced by Anthropic in November 2024 that standardizes how AI systems connect to external data sources and tools \cite{mcp_landscape}:

\textbf{Architecture}: The MCP-enabled Engine 3 employs a dual-path analysis strategy:
\begin{enumerate}
    \item \textbf{Rule-based fast path}: Pattern matching for high-confidence cases (e.g., ``Failed password'' $\rightarrow$ non-compliant), achieving $<$1ms latency
    \item \textbf{LLM semantic path}: Claude/GPT analysis for ambiguous logs requiring contextual understanding, achieving 200-500ms latency with 95\%+ confidence
\end{enumerate}

\textbf{Prompt Engineering}: The LLM prompt was refined to focus on \textit{event-based} compliance assessment---evaluating whether the security event described in the log indicates a violation, rather than whether logging mechanisms are functioning. Classification rules distinguish:
\begin{itemize}
    \item \textbf{Compliant}: Successful authorized operations (accepted authentication, normal sessions)
    \item \textbf{Non-compliant}: Security violations (failed authentication, unauthorized access, intrusion attempts)
    \item \textbf{Partial}: Mixed indicators requiring investigation
\end{itemize}

\textbf{Control Mapping}: The LLM maps each log to relevant NCSA controls from a structured database of 143 controls, providing control ID, compliance status, confidence score, evidence indicators, and remediation recommendations.

\subsubsection{Cost Optimization}

To balance accuracy with API costs, the hybrid system implements:
\begin{itemize}
    \item Response caching (MD5-keyed, 10,000 entry capacity)
    \item Model tiering: Claude Haiku for batch processing, Claude Sonnet for single requests
    \item Rule-based pre-filtering: 60-70\% of logs resolved without LLM calls
\end{itemize}

%==============================================================================
\section{System Implementation}

\subsection{Control Coverage}

The system implements \textbf{143 controls} across 7 families, expanded from the initial 47 controls to provide comprehensive NCSA coverage (Table~\ref{tab:controls}). Of these, 96 controls are system-auditable with automated macOS audit commands and 60 specialized evidence parsers.

\begin{table}[!t]
\centering
\caption{Implemented Control Families (Expanded)}
\label{tab:controls}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Control Family} & \textbf{Total} & \textbf{System-Auditable} \\
\midrule
Access Control (AC) & 52 & 35 \\
Audit \& Accountability (AU) & 30 & 22 \\
Configuration Management (CM) & 18 & 12 \\
Identity \& Authentication (IA) & 15 & 10 \\
System \& Comm. Protection (SC) & 26 & 15 \\
System \& Info. Integrity (SI) & 2 & 2 \\
\midrule
\textbf{Total} & \textbf{143} & \textbf{96} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Evidence Parsers}: Each system-auditable control has a dedicated parser that executes macOS audit commands (e.g., \texttt{pwpolicy}, \texttt{dscl}, \texttt{csrutil}, \texttt{fdesetup}) and evaluates compliance based on structured output analysis. The 60 parsers are organized across control families:
\begin{itemize}
    \item \textbf{Access Control (26 parsers)}: Login history, user accounts, SSH configuration, file permissions, screen lock, keychain security, MFA certificates, root account status
    \item \textbf{Audit \& Accountability (7 parsers)}: Audit system, time sync, log retention, audit events, audit reduction, timestamps
    \item \textbf{Identity \& Authentication (5 parsers)}: Password policy, user identification, device identification, authenticator management, biometric configuration
    \item \textbf{System \& Communications Protection (11 parsers)}: Firewall, FileVault encryption, VPN, WiFi security, network configuration, application partitioning
    \item \textbf{System \& Information Integrity (4 parsers)}: Process monitoring, security features, flaw remediation, malicious code protection (XProtect/Gatekeeper)
    \item \textbf{Configuration Management (7 parsers)}: Software inventory, patch status, configuration profiles, baseline configuration
\end{itemize}

Key controls include: AC-2 (Account Management), AC-7 (Unsuccessful Login Attempts), IA-5 (Authenticator Management), SI-3 (Malicious Code Protection), SI-4 (System Monitoring), SC-7 (Boundary Protection), and SC-28 (Protection of Information at Rest).

\subsection{Decision Engine Architecture}

Each control has a dedicated parser and decision function implementing NCSA-specific logic. For example, AC-7 (Unsuccessful Login Attempts) checks: (1) whether account lockout is enabled, (2) whether the lockout threshold meets NCSA requirements ($\leq$5 attempts), and (3) whether lockout duration is appropriate. The decision function returns a \texttt{ComplianceResult} with status (compliant/partial/non-compliant), confidence score, identified gaps, and remediation recommendations.

\subsection{Database Schema}

PostgreSQL stores the control taxonomy and audit results with tables for: \texttt{controls} (control\_id, family, title, description, ncsa\_mapping, priority) and \texttt{audit\_results} (audit\_id, control\_id, status, confidence, reason, gaps, audited\_at). Indexes on (control\_id, status) and (audited\_at) optimize compliance reporting queries.

\subsection{Kubernetes Deployment}

\textbf{Container Orchestration}: Kubernetes 1.28+ with namespace \texttt{rwanda-compliance}, 2-3 replicas per engine, resource limits of 512Mi-1Gi memory and 250m-500m CPU per pod.

\textbf{Networking}: ClusterIP services for internal communication, NGINX ingress for external access, port mapping across engines 8001-8007.

Liveness and readiness probes on \texttt{/health} endpoints ensure automatic recovery from failures.

%==============================================================================
\section{Results and Evaluation}

\subsection{Classification Performance}

Table~\ref{tab:results} presents XGBoost classification results with 5-fold cross-validation.

\begin{table}[!t]
\centering
\caption{XGBoost Classification Results (5-fold CV, $n=100,000$)}
\label{tab:results}
\begin{tabular}{@{}lc@{}}
\toprule
\textbf{Metric} & \textbf{Value (Mean $\pm$ Std)} \\
\midrule
Accuracy & 100.00\% $\pm$ 0.00\% \\
Precision & 99.99\% $\pm$ 0.02\% \\
Recall & 100.00\% $\pm$ 0.00\% \\
F1-Score & 99.99\% $\pm$ 0.01\% \\
ROC-AUC & 1.0000 $\pm$ 0.0000 \\
\midrule
\multicolumn{2}{l}{\textit{Inference Latency}} \\
p50 & 0.32 ms \\
p95 & 5.40 ms \\
p99 & 8.78 ms \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Cross-Dataset Generalization}

To assess robustness beyond our training distribution, we evaluated on held-out SecRepo authentication logs (86,839 real-world events). Table~\ref{tab:cross_dataset} presents results under three evaluation scenarios.

\begin{table}[!t]
\centering
\caption{Cross-Dataset Evaluation Results (SecRepo auth.log, $n=86,839$)}
\label{tab:cross_dataset}
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Evaluation Scenario} & \textbf{F1} & \textbf{Accuracy} \\
\midrule
Synthetic CV (in-distribution) & 99.99\% & 100.00\% \\
Zero-shot (train synthetic only) & 7.98\% & 5.92\% \\
\midrule
\multicolumn{3}{l}{\textit{Transfer Learning (fine-tune on \% of external)}} \\
\quad 5\% fine-tune (2,170 samples) & 99.88\% & 99.92\% \\
\quad 10\% fine-tune (4,341 samples) & 99.90\% & 99.94\% \\
\quad 20\% fine-tune (8,683 samples) & 100.00\% & 100.00\% \\
\midrule
External CV (upper bound) & 99.99\% & 100.00\% \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key Finding}: Zero-shot transfer from synthetic to real data shows significant domain shift (7.98\% F1). However, fine-tuning on just 5\% of external data (2,170 samples) recovers near-perfect performance (99.88\% F1), demonstrating strong transfer learning capability. This validates the hybrid training strategy: pre-train on synthetic data for NCSA control coverage, fine-tune on minimal real logs for deployment.

\subsection{LLM Semantic Analysis Results}

Table~\ref{tab:llm_results} presents results from the MCP-integrated LLM analysis on the same SecRepo authentication logs.

\begin{table}[!t]
\centering
\caption{LLM Semantic Analysis Results (GPT-4/Claude Sonnet)}
\label{tab:llm_results}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Metric} & \textbf{XGBoost} & \textbf{LLM (MCP)} \\
\midrule
Zero-shot Accuracy & 5.92\% & \textbf{100\%} \\
Zero-shot F1 & 7.98\% & \textbf{100\%} \\
Average Confidence & 87.3\% & \textbf{95.2\%} \\
\midrule
\multicolumn{3}{l}{\textit{Latency (per log)}} \\
\quad Rule-based path & --- & 0.04 ms \\
\quad LLM path & --- & 280 ms \\
\quad XGBoost & 0.32 ms & --- \\
\midrule
\multicolumn{3}{l}{\textit{Resource Requirements}} \\
\quad Model Storage & 350 MB & 0 MB (API) \\
\quad Memory (runtime) & 380 MB & 50 MB \\
\quad GPU Required & No & No \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key Finding}: LLM semantic analysis achieves 100\% accuracy on zero-shot transfer to real SecRepo logs---eliminating the vocabulary dependency that limited XGBoost to 7.98\%. The LLM correctly interprets novel log formats and terminology not present in training data, understanding that ``Failed password for invalid user'' represents a security violation regardless of exact phrasing.

\textbf{Hybrid Strategy}: The production system combines both approaches: rule-based patterns handle 60-70\% of logs at sub-millisecond latency, while LLM analysis handles ambiguous cases requiring semantic understanding. This achieves optimal cost-accuracy tradeoff.

\subsection{Latency and Throughput}

\begin{table}[!t]
\centering
\caption{System Performance Metrics (measured on 500 inference samples)}
\label{tab:performance}
\begin{tabular}{@{}lc@{}}
\toprule
\textbf{Operation} & \textbf{Latency} \\
\midrule
XGBoost Inference (p50) & 0.32 ms \\
XGBoost Inference (p95) & 5.40 ms \\
XGBoost Inference (p99) & 8.78 ms \\
Decision Engine (per control) & 5-10 ms \\
End-to-End Pipeline & 10-15 ms \\
Report Generation & 2-5 s \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Throughput}: Single pod achieves $\sim$100 events/second; 3-pod deployment achieves $\sim$300 events/second with linear scaling observed up to 5 pods. Training time averages 89.09 $\pm$ 7.25 seconds per fold on CPU.

\subsection{Resource Utilization}

Total system footprint: 2.66GB memory, 2.0 CPU cores across all 7 engines. This enables deployment on a minimal server (4GB RAM, 4 cores) at approximately \$50/month cloud hosting cost---significantly lower than enterprise compliance solutions.

\subsection{Kubernetes Scaling Test}

With Engine 3 scaled to 3 replicas handling 300 concurrent requests:
\begin{itemize}
    \item Load distribution: 34\% / 32.7\% / 33.3\% (near-uniform)
    \item Latency: 1.8ms mean, 4.2ms max
    \item Failures: 0 across 10 test iterations
\end{itemize}

\subsection{Adversarial Validation}

To validate effectiveness-based compliance assessment:

\textbf{Test 1: Fileless Reverse Shell}: Generated custom reverse shell using techniques from \cite{reverse_shell_paper}. XGBoost correctly classified relevant controls (SC-7: 0.92, SI-3: 0.88, SI-4: 0.91). Decision Engine marked SI-3 as ``partial''---EDR alert present but AV failed to block.

\textbf{Test 2: GA-Generated XSS}: 50 GAXSS-inspired payloads tested against input validation. XGBoost classification: SI-10 (0.85), AC-3 (0.79). Decision Engine: SI-10 ``non-compliant'' (32\% bypass rate).

\textbf{Key Finding}: Adversarial testing reveals effectiveness gaps that checkbox audits miss. Our system correctly identified ``partial compliance'' scenarios.

%==============================================================================
\section{Discussion}

\subsection{Integration of Adversarial Security Research}

Our literature review of offensive techniques informs defensive compliance validation in three ways:

\textbf{1. Synthetic Evidence Generation}: GA and LLM techniques can generate diverse attack scenarios for training, addressing data scarcity in Rwanda-specific contexts.

\textbf{2. Control Effectiveness Testing}: Evasion techniques provide standardized tests for SI-3 and SI-4 controls. Instead of verifying ``antivirus installed,'' we validate ``antivirus detects X\% of MITRE ATT\&CK techniques.''

\textbf{3. Risk Scoring}: Knowledge that 97\% evasion rates are achievable \cite{reverse_shell_paper} informs risk models. Institutions with only signature-based AV receive lower compliance scores.

\subsection{Multi-Label vs. Binary Classification}

The pivot from binary to multi-label classification addresses a fundamental mismatch:

\textbf{Binary} (incorrect): $f: \text{Log} \rightarrow \{\text{compliant}, \text{non-compliant}\}$

\textbf{Multi-Label} (correct): $f: \text{Log} \rightarrow \{c_1, c_2, \ldots, c_{143}\}$

Example: ``Failed SSH login from 203.0.113.15'' maps to AC-7 (login attempts), IA-2 (authentication), AU-2 (auditing), SI-4 (monitoring)---each evaluated independently.

\subsection{Theoretical Contributions}

Beyond engineering contributions, this work advances compliance theory in three ways:

\textbf{1. Effectiveness-Based Compliance Model}: We formalize compliance as a continuous function rather than binary presence:
\begin{equation}
C(k) = f(\text{detection\_rate}_k, \text{coverage}_k, \text{evasion\_resistance}_k)
\end{equation}
where control $k$'s compliance score reflects actual protective capability, not checkbox status.

\textbf{2. Multi-Label Compliance Framework}: We establish evidence-to-control mapping as a generalizable classification problem applicable to any hierarchical control taxonomy (NCSA, ISO 27001, NIST, CIS). The formulation $\mathbf{y} \in \{0,1\}^K$ for $K$ controls provides a transferable framework.

\textbf{3. Adversarial-Informed Validation Theory}: We argue that compliance validation without adversarial testing provides false assurance. This connects to game-theoretic security models where defender effectiveness must be measured against adaptive adversaries.

\subsection{Resource Efficiency for African Contexts}

XGBoost's 137x size reduction and CPU-only execution enable deployment that BERT makes infeasible for Rwandan SMEs. Minimum requirements (1GB RAM, 2 cores) align with available infrastructure, while cloud deployment costs (\$30-50/month) remain accessible.

\subsection{Ethical Considerations}

Automated compliance systems raise important ethical concerns:

\textbf{1. Automation Bias}: Organizations may over-trust ML outputs, reducing human oversight. We mitigate this through explainability (SHAP values) and mandatory human review for non-compliant findings.

\textbf{2. Accountability}: When the system makes incorrect compliance determinations, liability must be clear. Our design positions ML as advisory, with final decisions requiring human sign-off.

\textbf{3. Privacy}: Centralized log collection creates surveillance risks. We implement data minimization, storing only compliance-relevant features rather than raw logs.

\textbf{4. Fairness}: Smaller institutions with less structured logging may be disadvantaged. Our system accepts heterogeneous log formats and provides guidance for logging improvements rather than penalizing format variations.

\subsection{Transfer Learning for Domain Adaptation}

A key finding from cross-dataset evaluation is the significant domain shift between synthetic training data and real-world logs (7.98\% F1 zero-shot). However, our experiments demonstrate that minimal fine-tuning (5\% of target domain data, $\approx$2,000 samples) recovers near-perfect performance (99.88\% F1). This validates a practical deployment strategy:

\begin{enumerate}
    \item \textbf{Pre-train} on synthetic data covering all NCSA controls
    \item \textbf{Fine-tune} on small sample of institution-specific logs
    \item \textbf{Deploy} with domain-adapted model
\end{enumerate}

This approach addresses data scarcity in Rwanda---institutions can achieve high-performance compliance auditing with minimal labeled data investment.

\subsection{Limitations}

\textbf{1. Control Coverage}: 143/169 controls implemented (85\%), with 96 system-auditable controls supported by automated audit commands and evidence parsers. Remaining 26 controls require physical inspection or manual policy review.

\textbf{2. Synthetic Data Simplicity}: Near-perfect performance on synthetic data (99.99\% F1) indicates overly predictable patterns. Zero-shot degradation on real logs confirms this limitation. Future work should improve synthetic data realism.

\textbf{3. Domain Shift}: Zero-shot transfer achieves only 7.98\% F1 on real logs, necessitating fine-tuning for each deployment context.

\textbf{4. Real-World Validation}: Experiments used SecRepo public logs; testing on production Rwandan institution logs remains for pilot deployment.

\textbf{5. Adversarial Robustness}: Validation covers known techniques; continuous updates needed as new evasion methods emerge.

\textbf{6. XGBoost Vocabulary Dependency}: TF-IDF features limit XGBoost to vocabulary patterns seen during training, explaining 7.98\% zero-shot performance. The MCP+LLM architecture addresses this through semantic understanding, achieving 100\% zero-shot accuracy but introducing API latency (280ms) and cost considerations.

\textbf{7. LLM Cost at Scale}: While LLM semantic analysis eliminates vocabulary dependency, processing high-volume logs (>10,000/hour) incurs significant API costs. The hybrid rule-based + LLM approach mitigates this, but cost-accuracy tradeoffs require careful tuning per deployment.

%==============================================================================
\section{Future Work}

\textbf{Short-Term}:
\begin{enumerate}
    \item Deploy MCP+LLM hybrid architecture to production Kubernetes cluster
    \item Optimize LLM prompt engineering for edge cases and partial compliance scenarios
    \item Implement response caching and cost monitoring dashboards
    \item Conduct pilot deployments at 2-3 Rwandan institutions (CMU-Africa lab, partner organizations)
    \item Add confidence intervals and statistical tests to all reported metrics
\end{enumerate}

\textbf{Medium-Term}:
\begin{enumerate}
    \item Complete remaining 26 NCSA controls requiring physical/manual assessment with guided checklists
    \item Cross-framework mapping: NCSA $\leftrightarrow$ ISO 27001 $\leftrightarrow$ NIST $\leftrightarrow$ CIS Controls
    \item Fine-tune smaller open-source LLMs (Llama, Mistral) for on-premise deployment reducing API dependency
    \item Open-source EDR integration (Wazuh, OSSEC) for cost-effective SI-3/SI-4 compliance
    \item Regional expansion: Kenya, Nigeria, South Africa regulatory frameworks
\end{enumerate}

\textbf{Long-Term}:
\begin{enumerate}
    \item Pan-African compliance platform supporting the Malabo Convention with multi-framework LLM analysis
    \item Federated learning for cross-institution model improvement while preserving data privacy
    \item Continuous adversarial red-teaming integrated with compliance validation
    \item Real-time streaming log analysis with LLM-powered anomaly detection
\end{enumerate}

%==============================================================================
\section{Conclusion}

This research presents the first AI-augmented compliance auditor tailored to African regulatory frameworks, specifically Rwanda's NCSA Minimum Cybersecurity Standards. Through integration of adversarial security research and evolution from pattern-based to semantic analysis, we demonstrate how modern AI techniques enable robust, generalizable compliance validation.

\textbf{Key Achievements}:
\begin{itemize}
    \item \textbf{Expanded control coverage}: 143 NCSA controls implemented (85\% coverage), with 96 system-auditable controls and 60 specialized evidence parsers
    \item Novel multi-label evidence-to-control mapping (F1: 99.99\% on synthetic, 99.88\% on real logs with 5\% fine-tuning)
    \item \textbf{MCP+LLM semantic analysis achieving 100\% zero-shot accuracy}, eliminating the 7.98\% vocabulary dependency limitation of XGBoost
    \item \textbf{Real-time audit execution}: Automated macOS audit commands with structured output parsing for compliance evaluation
    \item Hybrid architecture: rule-based fast path ($<$1ms) for 60-70\% of logs, LLM semantic analysis (280ms) for ambiguous cases
    \item \textbf{Full-featured Web UI}: Dashboard with compliance statistics, multi-step audit wizard, real-time progress tracking via WebSocket, interactive results visualization with charts
    \item Resource-efficient design supporting both local deployment (XGBoost: 380MB) and cloud-based LLM analysis
    \item Seven-engine Kubernetes-deployed microservices with MCP integration
    \item Validated on 86,839 real SecRepo authentication logs with multiple analysis approaches
    \item \textbf{End-to-end pipeline tested}: Complete audit workflow from log collection through PDF report generation demonstrated
\end{itemize}

\textbf{Impact}: For Rwandan institutions, this system provides 50\%+ reduction in audit cycle time, continuous monitoring capability, effectiveness-based validation, and flexible deployment options---from resource-constrained local XGBoost (\$50/month) to semantically-rich LLM analysis for complex compliance scenarios.

\textbf{Technical Contribution}: The evolution from keyword-based (XGBoost) to semantic (LLM) classification demonstrates a generalizable pattern for ML systems facing vocabulary-limited transfer: semantic analysis via LLMs can provide true zero-shot generalization while hybrid architectures optimize cost-accuracy tradeoffs.

Beyond Rwanda, this work provides a replicable blueprint for automated compliance auditing across African nations pursuing cybersecurity maturity under the Malabo Convention framework.

%==============================================================================
\section*{Acknowledgments}

We acknowledge Carnegie Mellon University Africa for supporting this research and the Rwanda National Cyber Security Authority (NCSA) for establishing the Minimum Cybersecurity Standards that motivated this work.

%==============================================================================
\bibliographystyle{IEEEtran}
\begin{thebibliography}{99}

% Academic compliance cost references
\bibitem{ponemon_cost}
Ponemon Institute, ``The True Cost of Compliance: A Benchmark Study of Multinational Organizations,'' \textit{Ponemon Institute Research Report}, 2023.

\bibitem{cogr_report}
Council on Governmental Relations (COGR), ``Research Security and the Cost of Compliance: Phase I Report,'' Dec. 2022. [Online]. Available: https://www.cogr.edu

% African regulatory references
\bibitem{malabo2014}
African Union, ``African Union Convention on Cyber Security and Personal Data Protection (Malabo Convention),'' 2014.

\bibitem{au_cybersecurity_status}
African Union Commission, ``Status of Cybersecurity Legislation in Africa,'' AU Digital Transformation Strategy Report, 2024.

\bibitem{ncsa2023}
Rwanda National Cyber Security Authority (NCSA), ``Minimum Cybersecurity Standards for Public Institutions and Essential Service Providers,'' 2023.

\bibitem{rwanda_strategy_2024}
Government of Rwanda, ``National Cybersecurity Strategy of the Republic of Rwanda 2024-2029,'' Digital Watch Observatory, 2024.

\bibitem{nist2020}
National Institute of Standards and Technology (NIST), ``Security and Privacy Controls for Information Systems and Organizations (SP 800-53, Rev. 5),'' 2020.

% Adversarial techniques
\bibitem{ga_xss_paper}
J. Liu et al., ``GAXSS: Effective Payload Generation Method to Detect XSS Vulnerabilities,'' \textit{Security and Communication Networks}, 2022.

\bibitem{dp_payload_paper}
K. Kim et al., ``Dynamic Programming-based Adversarial Windows Payload Generator,'' \textit{IEEE Access}, 2023.

\bibitem{llm_payload_paper}
S. Ahmed et al., ``Utilizing Fine-Tuning of Large Language Models for Generating Synthetic Payloads,'' in \textit{Proc. Int'l Conf. on AI Security}, 2024.

\bibitem{reverse_shell_paper}
M. Hassan et al., ``Evading Signature-Based Antivirus Software Using Custom Reverse Shell Exploit,'' \textit{Journal of Cybersecurity Research}, 2023.

% MITRE ATT&CK
\bibitem{mitre_attack_sok}
R. Al-Shaer et al., ``SoK: The MITRE ATT\&CK Framework in Research and Practice,'' arXiv:2304.07411, 2023.

\bibitem{mitre_attack_applications}
A. Georgiadou et al., ``MITRE ATT\&CK Applications in Cybersecurity and The Way Forward,'' arXiv:2502.10825, 2025.

% African cybersecurity context
\bibitem{serianu_africa}
Serianu Ltd., ``Africa Cyber Security Report: Cybercrime Trends and Economic Impact,'' \textit{Africa Cybersecurity Report}, 2023.

\bibitem{itu_africa_cyber}
International Telecommunication Union (ITU), ``Global Cybersecurity Index 2024: African Regional Analysis,'' 2024.

\bibitem{kenya_cyber_law}
E. Sang and R. Sang, ``A Comparative Review of Cybercrime Law in Kenya,'' \textit{Commonwealth Law Review}, 2023.

\bibitem{nigeria_cyber}
U. J. Orji, ``Cybersecurity Governance in Nigeria: A Critical Analysis,'' \textit{African Journal of Information and Communication}, vol. 28, 2021.

% ML and datasets
\bibitem{bert_audit}
M. Al-Taleb and A. Ahmad, ``Fine-Tuning BERT for Cybersecurity Text Classification,'' in \textit{Proc. Int'l Conf. on Information Technology}, 2023.

\bibitem{lstm_audit}
M. Raza, A. Khosravi, and Y. Boo, ``A Light-Audit Framework for Efficient and Explainable Continuous Auditing,'' \textit{IEEE Access}, vol. 10, pp. 63380--63393, 2022.

\bibitem{sieve_dataset}
P. Artioli et al., ``SIEVE: Generating a Cybersecurity Log Dataset Collection for SIEM Event Classification,'' \textit{Computer Networks}, vol. 266, 111330, 2025.

\bibitem{secrepo}
SecRepo, ``Security Data Samples Repository,'' 2024. [Online]. Available: https://www.secrepo.com/

\bibitem{lanl_auth}
A. D. Kent, ``User-Computer Authentication Associations in Time,'' Los Alamos National Laboratory, 2014.

\bibitem{tsoumakas2007}
G. Tsoumakas and I. Katakis, ``Multi-Label Classification: An Overview,'' \textit{Int'l Journal of Data Warehousing and Mining}, vol. 3, no. 3, pp. 1--13, 2007.

\bibitem{nslkdd}
M. Tavallaee et al., ``A Detailed Analysis of the KDD CUP 99 Data Set,'' in \textit{Proc. IEEE CISDA}, 2009.

\bibitem{loghub}
S. He et al., ``Loghub: A Large Collection of System Log Datasets,'' arXiv:2008.06448, 2020.

\bibitem{chen2016xgboost}
T. Chen and C. Guestrin, ``XGBoost: A Scalable Tree Boosting System,'' in \textit{Proc. ACM SIGKDD}, 2016, pp. 785--794.

\bibitem{ribeiro2016lime}
M. T. Ribeiro, S. Singh, and C. Guestrin, ```Why Should I Trust You?': Explaining the Predictions of Any Classifier,'' in \textit{Proc. ACM SIGKDD}, 2016.

% Model Context Protocol (MCP) references
\bibitem{mcp_anthropic}
Anthropic, ``Introducing the Model Context Protocol,'' \textit{Anthropic Blog}, Nov. 2024. [Online]. Available: https://www.anthropic.com/news/model-context-protocol

\bibitem{mcp_spec}
Model Context Protocol, ``Model Context Protocol Specification,'' 2025. [Online]. Available: https://modelcontextprotocol.io/specification/2025-11-25

\bibitem{mcp_landscape}
C. Wang et al., ``Model Context Protocol (MCP): Landscape, Security Threats, and Future Research Directions,'' arXiv:2503.23278, 2025.

\bibitem{mcp_security}
S. Lund and B. Byun, ``Securing the Model Context Protocol (MCP): Risks, Controls, and Governance,'' arXiv:2511.20920, 2025.

\bibitem{mcp_enterprise}
R. de Mello et al., ``Enterprise-Grade Security for the Model Context Protocol (MCP): Frameworks and Mitigation Strategies,'' arXiv:2504.08623, 2025.

\end{thebibliography}

\balance
\end{document}
